---
title: "Day 1 Part 3.1 ML 4 DH 2019"
author: "Dave Campbell, dac5@sfu.ca, @iamdavecampbell"
date: '2019-06-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


In this part:

- Showing a larger webscraping routine

### Step 5: Let's get all of Bob Dylan's lyrics!
Here we go to https://www.bobdylan.com/songs/ and find all of the links to songs.  Then we visit each song link and extract its lyrics.  We do this by first finding links and then creating a _for loop_ that will allow us to go through all links one at a time.  Note that webscraping is fragile, often breaks, and stops working when a webpage is updated in a new style.


The process is simple but takes a lot of trial an error.  We need to look at the original webpage.  Using the inspector or the _selectorgadget_ from the _rvest_ library to help us figure out where in the html code the useful information is stored.  Then we target these specific pieces.  

Using piping makes it easy to focus in on a chunk of html code at a time.  It also allows us to run pieces and build on code until we have what we need.



```{r, eval = FALSE}
SongWebpage = read_html("https://www.bobdylan.com/songs/")
Allsongs = SongWebpage %>% 
             html_nodes("div.line.line_detail") 
  

#Make a place to put everything:
AllBob = tibble(title =  NA, credit = NA, lyrics = NA, album = NA,.rows = length(Allsongs))


for(song in 1:length(Allsongs)){
  link = Allsongs[song]%>%
        html_nodes("span.song")%>% 
        html_nodes("a") %>% 
        html_attr("href")  
  #Sys.sleep(rpois(1,1))
  Webpage = read_html(link)
    AllBob$title[song] = Webpage %>% 
      html_nodes( "h2.headline") %>%
      html_text()
    AllBob$credit[song] = Webpage %>% 
      html_nodes( "div.credit") %>%
      html_text()
    AllBob$lyrics[song] = Webpage %>% 
      html_nodes( "div.article-content.lyrics") %>%
      html_text()
    AllBob$album[song] = (Webpage %>% 
      html_nodes( "div.information") %>%
      html_text())[1]
}

### Some data cleaning; to get rid of formatting
# extra \t white space in the lyrics
AllBob$lyrics = gsub(AllBob$lyrics,pattern="(\\t)+",replacement = "")
# extra white space (\s) and formatted white space (\t), line breaks (\n) and the words 'Written by:'
AllBob$credit = gsub(AllBob$credit,pattern="(^\\s+)|(\\s+$)|(\\t+)|(\\n)+|(Written by: )",replacement = "")
# extra \t white space and line breaks \n in the album name.  We also don't care about the words '(Original Release)'
AllBob$album = gsub(AllBob$album,pattern="(\\t+)|(\\n)+|(\(Original Release\))",replacement = "")


save("AllBob",file = "AllBobLyrics.Rds")


AllBobTokenWords  = unnest_tokens(AllBob,output = word,input = lyrics, token = "words")

save("AllBobTokenWords",file = "AllBobTokenWords.Rds")

```


